A sheaf--function dictionary for algebraic tori over local fields

### 2013-08-19 shanghai.tex
Moved the lemma before the proposition on restriction to the identity compnent. Added proof that G^0 is an algebraic group.
Some other small improvements.

To define the category of finite quasicharacter sheaves we must work a bit harder.
Consider the following category $C(G)$.
Let us say that $f : H\to G$ is a \emph{discrete isogeny} if it is a a finite, surjective
\'etale morphism and a morphism of commutative group $\Fq$-schemes, for which
the action of $\Gal(\bFq/\Fq)$ on the \'etale group scheme $f^{-1}{\bar e}$ is trivial.
An object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ is a discrete isogeny 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\alpha,T)$ 
where $\alpha$ is a morphism of schemes such that $f' = f\circ \alpha$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker '$ on $V$ by $\psi \circ \alpha$.
\[
\begin{tikzcd}
V \arrow{d}{T} & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} & \arrow[equal]{d} G\\
V' & \Aut(V) & \arrow[swap]{l}{\psi} \arrow[swap]{u}{\alpha\vert_{\ker f}} \ker f \arrow[hook]{r} & \arrow[swap]{u}{\alpha} H' \arrow{r}{f'} & G
\end{tikzcd}
\]
We say that $(\alpha,T)$ in $C(G)$ is a \emph{weak isomorphism} if $\alpha$ is surjective
and $T$ is an isomorphism of vector spaces.
Let \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at weak isomorphisms.

@@

\begin{proof}
  The identity component $G^0$ of $G$ is a connected algebraic group over $\Fq$, 
  \Clifton{This needs to be treated more carefully!}
  clopen in $G$ \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}. 
    By Proposition~\ref{prop:connected}, every
  quasicharacter sheaf on $G^0$ is isomorphic to a 
  finite quasicharacter sheaf on $G^0$, 
  so to prove the proposition it suffices to show that every 
  finite quasicharacter sheaf on $G^0$ extends to a quasicharacter sheaf on $G$.

\begin{lemma}\label{lemma:ext}
Every discrete isogeny (see Section~\ref{sec:finite}) to $G^0$ extends to a discrete
isogeny to $G$ such that the component group of the covering space is isomorphic
to the component group of $G$.
\end{lemma}

@@

@@

The techniques in this paper apply to all non-archimdean local fields $K$ with residue field $k$,
and to all algebraic tori $T$ over $K$, without placing any restrictions on the characteristic of $K$
or on the ramification of $T$. Because of this, it is natural to ask if we can compare
quasicharacter sheaves on $T$ and on $T'$ under conditions when there is a
natural comparison between characters of $T(K)$ and of $T'(K')$.
This comparison of has been made precise in \cite{chai-yu:01a}, \David{more specific citation?}
building on ideas of \cite{deligne:84a}.

>>>>

To define the category of finite quasicharacter sheaves we must work a bit harder.
Consider the following category $C(G)$.
Let us say that $f : H\to G$ is a \emph{discrete isogeny} if it is a a finite, surjective
\'etale morphism and a morphism of commutative group $\Fq$-schemes, for which
the action of $\Gal(\bFq/\Fq)$ on the \'etale group scheme $f^{-1}({\bar e})$ is trivial.
An object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ is a discrete isogeny 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\alpha,T)$ 
where $\alpha$ is a morphism of schemes such that $f' = f\circ \alpha$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker '$ on $V$ by $\psi \circ \alpha$.
\[
\begin{tikzcd}
V \arrow{d}{T} & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} & \arrow[equal]{d} G\\
V' & \Aut(V) & \arrow[swap]{l}{\psi} \arrow[swap]{u}{\alpha\vert_{\ker f}} \ker f \arrow[hook]{r} & \arrow[swap]{u}{\alpha} H' \arrow{r}{f'} & G
\end{tikzcd}
\]
We say that $(\alpha,T)$ in $C(G)$ is a \emph{weak isomorphism} if $\alpha$ is surjective
and $T$ is an isomorphism of vector spaces.
Let \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at weak isomorphisms.

@@

Recall from Section~\ref{sec:finite} that 
a morphism $f : H\to G$ of group $\Fq$-schemes is a discrete isogeny 
if it is a finite, surjective \'etale morphism and 
the action of $\Gal(\bFq/\Fq)$ on the \'etale group scheme $f^{-1}{\bar e}$ is trivial.

\begin{lemma}\label{lemma:ext}
Every discrete isogeny  to $G^0$ extends to a discrete
isogeny to $G$ such that the component group of the covering space is isomorphic
to the component group of $G$.
\end{lemma}

@@

\begin{proof}
  Since $G$ is a smooth group scheme locally of finite type over $\Fq$, its
  identity component $G^0$ of $G$ is a connected smooth, 
  geometrically reduced group scheme of finite type over $\Fq$, by
   \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}. 
	Since $\Fq$ is a finite field, $G^0$ is actually reduced over $\Fq$;
	see \cite{stacks-project}*{\href{http://stacks.math.columbia.edu/tag/035X}{Tag 035X}}.
	It follows that $G^0$ is a connected algebraic group.
    By Proposition~\ref{prop:connected}, every
  quasicharacter sheaf on $G^0$ is isomorphic to a 
  finite quasicharacter sheaf on $G^0$, 
  so to prove the proposition it suffices to show that every 
  finite quasicharacter sheaf on $G^0$ extends to a quasicharacter sheaf on $G$.

@@

The techniques in this paper apply to all non-archimdean local fields $K$ with residue field $k$,
and to all algebraic tori $T$ over $K$, without placing any restrictions on the characteristic of $K$
or on the ramification of $T$. Because of this, it is natural to ask if we can compare
quasicharacter sheaves on $T$ and on $T'$ under conditions when there is a
natural comparison between characters of $T(K)$ and of $T'(K')$.
This comparison of has been made precise in \cite{chai-yu:01a}, \David{more specific citation?}\Clifton{In this case, maybe not necessary.}
building on ideas of \cite{deligne:84a}.
